---
title: "TexasCovid README"
author: "Nikolai Priser"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(knitr)
require(kableExtra)

```
## Summary 
Daily publication of cleaned and tidy Texas county-level Covid-19 statistics, as published by Texas DSHS.\
Original data sourced from <https://www.dshs.state.tx.us/coronavirus/additionaldata/>; ugly excel, beware. 

Tidy data can be accessed here:

* Daily Cases: <https://raw.githubusercontent.com/nikolkj/Texas-Covid/master/daily-county-data/Texas-County-Cases.csv>
* Daily Fatalities: <https://raw.githubusercontent.com/nikolkj/Texas-Covid/master/daily-county-data/Texas-County-Deaths.csv>
* Daily Tests: <https://raw.githubusercontent.com/nikolkj/Texas-Covid/master/daily-county-data/Texas-County-Tests.csv>

Data has been cleaned at put in a *long* format for easy visualization and modeling.

All data-tables have the following fields:

1. "County": Texas county name <factor>
2. "Date": Date associated with observation, YYYY-MM-DD format. 
3. "DailyCount": Aggregate measure, to-date, as published by DSHS.
4. "DailyDelta": Calculated daily measure ($x_{t} - x_{t-1}$) to get e.g. new cases for a given day
5. "LastUpdateDate": Date when the data was pulled.

DSHS updates data everyday around ~9:30am CST, tidy-data is then updated at 10:30am CST. 
&nbsp;

----
## Getting Data

Read data from github link.
```{r, echo =TRUE}
dat = read_csv(file = "https://raw.githubusercontent.com/nikolkj/Texas-Covid/master/daily-county-data/Texas-County-Cases.csv", col_names = TRUE, progress = FALSE)
```

Examine some data sample.
```{r, echo =TRUE}
dat %>% 
  filter(Date > "2020-04-15", DailyCount > 100) %>%
  sample_n(15) %>% 
  kable() %>% kableExtra::kable_styling(kable_input = ., bootstrap_options = c("striped", "hover"))

```

----
## Reporting Data

Find when new cases peaked for each county, take top 10.
```{r, echo=TRUE, fig.align='left'}
dat %>% group_by(County) %>%
  filter(DailyDelta == max(DailyDelta, na.rm = T)) %>%
  rename(PeakDate = Date, PeakCases = DailyDelta) %>%
  arrange(desc(PeakCases)) %>% head(n = 10) %>% 
  select(County, PeakDate, PeakCases) %>%
  kable() %>% kableExtra::kable_styling(kable_input = ., bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "left")

```

----
## Plotting Data

```{r, echo=TRUE, fig.fullwidth = TRUE}
dat %>%
  filter(!is.na(DailyDelta), 
         County %in% c("Harris","Dallas","Bexar","Walker")) %>%
  mutate(County = factor(County)) %>%
  select(County, Date, DailyDelta) %>% 
  ggplot(data = ., mapping = aes(x = Date, y = DailyDelta, col = County)) +
  geom_line() + 
  ggtitle("New Cases", subtitle = "For select counties") +
  ylab("") + xlab("") +
  scale_x_date(labels = scales::date_format(format = "%m/%d")) + 
  ggthemes::theme_fivethirtyeight()

```


```{r, echo=TRUE, fig.align='left', fig.fullwidth = TRUE}
dat %>% 
  filter(County %in% c("Harris","Dallas","Bexar","Walker"),
         DailyCount > 0,
         Date > "2020-03-15") %>%
  mutate(County = factor(County)) %>%
  select(County, Date, DailyCount) %>% 
  ggplot(data = ., mapping = aes(x = Date, y = DailyCount, col = County)) +
  geom_line() + 
  ggtitle("Total Cases", subtitle = "For select counties") +
  ylab("") + xlab("") +
  scale_y_continuous(na.value = 0, trans = "log10", labels = scales::number_format(big.mark = ",", accuracy = 1)) +
  scale_x_date(labels = scales::date_format(format = "%m/%d")) + 
  ggthemes::theme_fivethirtyeight()
  
```


&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;

